---
title: "Lab3A"
author: "Anastasia Galperina"
date: "9/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
  library(tidyverse)
```

## Download data into data folder
```{r, eval = FALSE}
   download.file(url="https://ndownloader.figshare.com/files/2292169",
              destfile = "data/portal_data_joined.csv")
```

```{r}
  surveys <- read_csv("data/portal_data_joined.csv")
  head(surveys)
```

# Challenge 1
```{r}
  str(surveys)
```
What is the class of the object surveys?

  Class: Data Frame
  
How many rows and how many columns are in this object?

  Rows: 34786 ; Columns: 13
  
How many species have been recorded during these surveys?

  Species: 48
  
# Indexing and Subsetting
```{r, results = "hide"}
  # first element in the first column of the data frame (as a vector)
  surveys[1, 1]   
  # first element in the 6th column (as a vector)
  surveys[1, 6]   
  # first column of the data frame (as a vector)
  surveys[, 1]    
  # first column of the data frame (as a data.frame)
  surveys[1]      
  # first three elements in the 7th column (as a vector)
  surveys[1:3, 7] 
  # the 3rd row of the data frame (as a data.frame)
  surveys[3, ]    
  # equivalent to head_surveys <- head(surveys)
  head_surveys <- surveys[1:6, ] 
  # The whole data frame, except the first column
  surveys[, -1]        
  # Equivalent to head(surveys)
  surveys[-(7:34786), ] 
  
  surveys["species_id"]       # Result is a data.frame
  surveys[, "species_id"]     # Result is a vector
  surveys[["species_id"]]     # Result is a vector
  surveys$species_id          # Result is a vector

```
  
  Results above are set to hidden
  
# Challenge 2
```{r}
  # Question 1
    surveys_200 <- surveys[200, ]
  # Question 2
    tail(surveys)

    n_rows <- nrow(surveys)
    surveys_last <- surveys[n_rows, ]
    surveys_last
  # Question 3
    surveys_middle <- surveys[n_rows / 2, ]
    surveys_middle
  # Question 4
    surveys_head <- surveys[-(7:n_rows), ]
    surveys_head
```

# Factors
```{r}
  sex <- factor(c("male", "female", "female", "male"))
```

```{r}
   levels(sex)
   nlevels(sex)
```

```{r}
  sex
```

```{r}
  sex <- factor(sex, levels = c("male", "female"))
  sex
```

```{r}
  as.character(sex)
```
  
```{r, eval = FALSE}
  year_fct <- factor(c(1990, 1983, 1977, 1998, 1990))
  as.numeric(levels(year_fct))[year_fct]    # The recommended way.
```

```{r}
  plot(as.factor(surveys$sex))
```

```{r}
   sex <- factor(surveys$sex)
   head(sex)
```

```{r}
  levels(sex)
```

```{r}
  levels(sex)[1] <- "undetermined"
  levels(sex)
```

```{r}
  head(sex)
```

# Challenge 3
```{r}
  levels(sex)[2:3] <- c("female", "male")
  sex <- factor(sex, levels = c("female", "male", "undetermined"))
  plot(sex)
```

```{r}
## Compare the difference between our data read as `factor` vs `character`.
surveys <- read.csv("data/portal_data_joined.csv", stringsAsFactors = TRUE)
str(surveys)
surveys <- read.csv("data/portal_data_joined.csv", stringsAsFactors = FALSE)
str(surveys)
## Convert the column "plot_type" into a factor
surveys$plot_type <- factor(surveys$plot_type)
```

# Challenge 4
```{r}
      animal_data <- data.frame(
              animal = c("dog", "cat", "sea cucumber", "sea urchin"),
              feel = c("furry","furry", "squishy", "spiny"),
              weighlabt = c(45, 8, 1.1, 0.8)
              )
    
      country_climate <- data.frame(
           country = c("Canada", "Panama", "South Africa", "Australia"),
           climate = c("cold", "hot", "temperate", "hot/temperate"),
           temperature = c(10, 30, 18, 15),
           northern_hemisphere = c(TRUE, TRUE, FALSE, FALSE),
           has_kangaroo = c(FALSE, FALSE, FALSE, TRUE)
           )
```
Can you predict the class for each of the columns in the following example? Check your guesses using str(country_climate):
Are they what you expected? Why? Why not?

  country, climate, temperature, and northern_hemisphere are factors; has_kangaroo is
  numeric
  
What would have been different if we had added stringsAsFactors = FALSE when creating the data frame?

   using stringsAsFactors = FALSE would have made character vectors instead of 
   factors
   
What would you need to change to ensure that each column had the accurate data type?

  removing the quotes in temperature and northern_hemisphere and replacing 1 by TRUE   in the has_kangaroo column would give what was probably intended

```{r}
  library("lubridate")
```

# Formatting Dates
```{r}
  my_date <- ymd("2015-01-01")
  str(my_date)
  my_date <- ymd(paste("2015", "1", "1", sep = "-")) 
  str(my_date)
  head(paste(surveys$year, surveys$month, surveys$day, sep = "-"))
  head(ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-")))
  head(surveys$date <- ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-")))
  str(surveys)
  
  summary(surveys$date)
  missing_dates <- surveys[is.na(surveys$date), c("year", "month", "day")]
  head(missing_dates)
```

