---
title: "Lab3B"
author: "Anastasia Galperina"
date: "9/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
  library(tidyverse)
```

## Download data into data folder
```{r, eval = FALSE}
   download.file(url="https://ndownloader.figshare.com/files/2292169",
              destfile = "data/portal_data_joined.csv")
```

```{r}
  surveys <- read_csv("data/portal_data_joined.csv")
  head(surveys)
```

# Selection and Filtering
```{r}
  select(surveys, plot_id, species_id, weight)
  select(surveys, -record_id, -species_id)
  filter(surveys, year == 1995)
```

# Pipes
```{r}
  surveys2    <- filter(surveys, weight < 5)
  surveys_sml <- select(surveys2, species_id, sex, weight)
  surveys_sml <- select(filter(surveys, weight < 5), species_id, sex, weight)

  surveys %>%
    filter(weight < 5) %>%
    select(species_id, sex, weight)
  
  surveys_sml <- surveys %>%
    filter(weight < 5) %>%
    select(species_id, sex, weight)

  surveys_sml
```

# Challenge 1
```{r}
  surveys %>%
    filter(year < 1995) %>%
    select(year, sex, weight)
```

# Mutate
```{r}
  surveys %>%
    mutate(weight_kg = weight / 1000)
  surveys %>%
    mutate(weight_kg = weight / 1000,
      weight_lb = weight_kg * 2.2)
  surveys %>%
    mutate(weight_kg = weight / 1000) %>%
    head()
  surveys %>%
    filter(!is.na(weight)) %>%
    mutate(weight_kg = weight / 1000) %>%
    head()
```

# Challenge 2
```{r}
  surveys_hindfoot_cm <- surveys %>%
    filter(!is.na(hindfoot_length)) %>%
    mutate(hindfoot_cm = hindfoot_length / 10) %>%
    filter(hindfoot_cm < 3) %>%
    select(species_id, hindfoot_cm)
```

# Split-Apply-Combine & Summarize()
```{r}
  surveys %>%
    group_by(sex) %>%
    summarize(mean_weight = mean(weight, na.rm = TRUE))
  surveys %>%
    group_by(sex, species_id) %>%
    summarize(mean_weight = mean(weight, na.rm = TRUE)) %>% 
    tail()
  surveys %>%
    filter(!is.na(weight)) %>%
    group_by(sex, species_id) %>%
    summarize(mean_weight = mean(weight))
  surveys %>%
    filter(!is.na(weight)) %>%
    group_by(sex, species_id) %>%
    summarize(mean_weight = mean(weight)) %>%
    print(n = 15)
  surveys %>%
    filter(!is.na(weight)) %>%
    group_by(sex, species_id) %>%
    summarize(mean_weight = mean(weight),
            min_weight = min(weight))
  surveys %>%
    filter(!is.na(weight)) %>%
    group_by(sex, species_id) %>%
    summarize(mean_weight = mean(weight),
            min_weight = min(weight)) %>%
    arrange(min_weight)
  surveys %>%
    filter(!is.na(weight)) %>%
    group_by(sex, species_id) %>%
    summarize(mean_weight = mean(weight),
            min_weight = min(weight)) %>%
    arrange(desc(mean_weight))

```

# Counting
```{r}
  surveys %>%
    count(sex) 
  surveys %>%             
    group_by(sex) %>%
    summarise(count = n())
  surveys %>%
    count(sex, sort = TRUE) 
  surveys %>%
    count(sex, species) 
  surveys %>%
    count(sex, species) %>%
    arrange(species, desc(n))
```

# Challenge 2

How many animals were caught in each plot_type surveyed?
```{r}
  surveys %>%
    count(plot_type) 
```
Use group_by() and summarize() to find the mean, min, and max hindfoot length for each species 
```{r}
  surveys %>%
    filter(!is.na(hindfoot_length)) %>%
    group_by(species_id) %>%
    summarize(
        mean_hindfoot_length = mean(hindfoot_length),
        min_hindfoot_length = min(hindfoot_length),
        max_hindfoot_length = max(hindfoot_length),
        n = n()
    )
```
What was the heaviest animal measured in each year? 
```{r}
  surveys %>%
    filter(!is.na(weight)) %>%
    group_by(year) %>%
    filter(weight == max(weight)) %>%
    select(year, genus, species, weight) %>%
    arrange(year)

```

# Gather and Spread
```{r}
  surveys_gw <- surveys %>%
    filter(!is.na(weight)) %>%
    group_by(plot_id, genus) %>%
    summarize(mean_weight = mean(weight))
  str(surveys_gw)

  surveys_spread <- surveys_gw %>%
    spread(key = genus, value = mean_weight)
  str(surveys_spread)

  surveys_gw %>%
    spread(genus, mean_weight, fill = 0) %>%
    head()
  
  surveys_gather <- surveys_spread %>%
    gather(key = "genus", value = "mean_weight", -plot_id)
  str(surveys_gather)
  
  surveys_spread %>%
    gather(key = "genus", value = "mean_weight", Baiomys:Spermophilus) %>%
    head()
```

# Challenge 3

Spread the surveys data frame with year as columns, plot_id as rows, and the number of genera per plot as the values. 
```{r}
  surveys_spread_genera <- surveys %>%
    group_by(plot_id, year) %>%
    summarize(n_genera = n_distinct(genus)) %>%
    spread(year, n_genera)
  head(surveys_spread_genera)


```
Now take that data frame and gather() it again, so each row is a unique plot_id by year combination.
```{r}
  surveys_spread_genera %>%
    gather("year", "n_genera", -plot_id)
```
Use gather() to create a dataset where we have a key column called measurement and a value column that takes on the value of either hindfoot_length or weight
```{r}
  surveys_long <- surveys %>%
    gather("measurement", "value", hindfoot_length, weight)
```
Then spread() them into a data set with a column for hindfoot_length and weight
```{r}
  surveys_long %>%
    group_by(year, measurement, plot_type) %>%
    summarize(mean_value = mean(value, na.rm=TRUE)) %>%
    spread(measurement, mean_value)
```

# Exporting Data
```{r}
  surveys_complete <- surveys %>%
  filter(!is.na(weight),           
         !is.na(hindfoot_length),  
         !is.na(sex))                

  ## Extract the most common species_id
  species_counts <- surveys_complete %>%
    count(species_id) %>% 
    filter(n >= 50)

  ## Only keep the most common species
  surveys_complete <- surveys_complete %>%
    filter(species_id %in% species_counts$species_id)

```
