---
title: "Lab4B"
author: "Anastasia Galperina"
date: "9/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
  library(tidyverse)
```

```{r}
  surveys_complete <- read_csv("data/surveys_complete.csv")
```

# Plotting
```{r}
  ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
  geom_point()
```

```{r}
  # Assign plot to a variable
  surveys_plot <- ggplot(data = surveys_complete, 
                       mapping = aes(x = weight, y = hindfoot_length))

  # Draw the plot (+ must be in first line to function)
  surveys_plot + 
    geom_point()
```

# Challenge 1

```{r}
  library(hexbin)
```

```{r}
  surveys_plot +
  geom_hex()
```

What are the relative strengths and weaknesses of a hexagonal bin plot compared to a scatter plot? 

  Hexbin seems to show more of a range while a scatter plot has the exact data points visible. 
  I think that a scatter plot would be more useful for plotting precise data while hexbin might
  be useful to plot large data sets where the data is really scattered to be able to compare 
  different data ranges.
  
# Building Plots Iteratively

```{r}
  ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
    geom_point()
```

```{r}
  ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1)
```

```{r}
  ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1, color = "blue")
```

```{r}
ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1, aes(color = species_id))
```

# Challenge 2

```{r}
  ggplot(data = surveys_complete, 
       mapping = aes(x = species_id, y = weight)) +
       geom_point(aes(color = plot_type))
```

# Boxplot

```{r}
  ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_boxplot()
```

```{r}
  ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_boxplot(alpha = 0) +
    geom_jitter(alpha = 0.3, color = "tomato")
```

```{r}
  ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_boxplot(alpha = 1) +
    geom_jitter(alpha = 0.01, color = "tomato")
```

# Challenge 3

```{r}
  ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_violin() +
    scale_y_log10()
```

```{r}
  ggplot(data = surveys_complete, mapping = aes(x = species_id, y = hindfoot_length)) +
    geom_boxplot(color = "red") + 
    geom_jitter(alpha = 0.01, color = "blue")
```

