---
title: "Lab9"
author: "Anastasia Galperina"
date: "10/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(stringr)
library(tidyverse)
```

# 14.3.1 Basic matches
```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
str_view_all(x, "an")
```

```{r}
str_view(x, ".a.")
str_view(x, "a.") # Should also find 'a' i apply
```

```{r}
# To create the regular expression, we need \\
dot <- "\\."

# But the expression itself only contains one:
writeLines(dot)
```

```{r}
# And this tells R to look for an explicit .
x <- c("abc", "a.c", "bef")
str_view(x, "a\\.c")
```

```{r}
x <- "a\\b"
writeLines(x)

str_view(x, "\\\\")
```

# Exercise
1. 
  ""\"" is used to as an escape character in regular expressions.
  "\\" creates the regular expression "\" and escapes that regular expression.
  "\\\" is used to as an escape character in regular expression "\\" and escapes that regular expression.
  
# 14.3.2 Anchors
```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a") # Matches first letter of string
```

```{r}
str_view(x, "a$") # Matches last letter of string
```

```{r}
# Will mark first occurance of apple in a string
x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")
```

```{r}
str_view(x, "^apple$")
```

```{r}
x <- c("apple pie", "apple apple", "apple cake")
y <- c("apple pie", "good apple", "apple cake")
z <- c("apple pie", "good apples", "apple cake")
str_view(x, "^apple$")
str_view(y, "^apple$")
str_view(z, "^apple$")
```

# Exercise
```{r}
x <- c("apples, pears, and bananas")
str_split(x, ", +(and +)?")[[1]]
```

# 14.3.3 Character classes and alternatives
```{r}
# Look for a literal character that normally has special meaning in a regex
x <- c("abc", "a.c", "a*c", "a c")
str_view(x, "a[.]c")
str_view(x, ".[*]c")
str_view(x, "a[ ]")

y <- c("abc", "a.c", "a*c", "a c", "b.c")
str_view(y, ".[*]c")
str_view(x, "a[*]c")
str_view(x, "a[//s]c")
```

```{r}
x <- c("grey", "gray")
str_view(x, "gr(e|a)y")
```

#Exercise
```{r}
x <- c("grey", "gray","apple","egg", "aeio","red","reed","reading","realise")
str_view(x, "^[aeiou]")
str_view_all(x, "[^aeiou]")
str_view(x, "[^e]ed$")
str_view(x, "(ing|ise)$")
```

#14.3.4 Repetition
```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")
str_view(x, "CC+")
str_view(x, 'C[LX]+')
str_view(x, "C{2}")
str_view(x, "C{2,}")
str_view(x, 'C[LX]+?') # ? gets the last occurance

y <- "1888 is the longest year in Roman numerals: MDCCCLLLLXXXVIII"
str_view(y, 'C[LX]+')
```

#Exercise
1.
  '?' matches a single occurrence (the last occurrence) of an expression: {0,1}
  '+' matches 1 or more occurences: {1,}
  '*' matches 0 or more occurences: {0,}
  
# 14.3.5 Grouping and backreferences
```{r}
fruit <- c("banana", "coconut", "cucumber", "apple", "orange")
str_view(fruit, "(..)\\1", match = TRUE)
```

# Exercise
1.
  (.)\1\1              : same character appearing three times in a row
  (.)(.)\\2\\1         : set of 2 characters followed by the same set of characters in reversed order
  (..)\1               : set of 2 characters repeated twice
  (.).\\1.\\1          : a character repeated followed by any random charcter in between repetitions
  (.)(.)(.).*\\3\\2\\1 : set of 3 characters followed by any character then the set of 3 characters in reversed order
  
# 14.4 Tools
```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")
```

```{r}
# How many common words start with t?
sum(str_detect(words, "^t")) # or just 'word'?
```

```{r}
# What proportion of common words end with a vowel?
mean(str_detect(words, "[aeiou]$")) 

```

```{r}
# Find all words containing at least one vowel, and negate
no_vowels_1 <- !str_detect(words, "[aeiou]")

# Find all words consisting only of consonants (non-vowels)
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)
```

```{r}
words[str_detect(words, "x$")]
str_subset(words, "x$")
```

```{r}
df <- tibble(
  word = words, 
  i = seq_along(word)
)
df %>% 
  filter(str_detect(word, "x$"))
```

```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")
str_count(x, "p")

y <- c("apple", "banana", "pear", "aaaaaaa")
str_count(y, "a")
```

```{r}
# On average, how many vowels per word?
mean(str_count(words, "[aeiou]"))
```

```{r}
df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )
```

```{r}
# Notice no overlap
str_count("abababa", "aba")
str_view_all("abababa", "aba")
```

# Exercise
```{r}
words[str_detect(words, "^x|x$")]
words[str_detect(words, "^[aeiou]|[^aeiou]$")]
 
str_subset(words, "^[aeiou].*[^aeiou]$") 
```

# 14.4.2 Extract matches
```{r}
length(sentences)
```

```{r}
head(sentences)
```

```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|") # Covert to regex
colour_match

has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match) # Only extracts first match
head(matches)

more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all

str_extract(more, colour_match)

str_extract_all(more, colour_match) # Returns list
```

```{r}
str_extract_all(more, colour_match, simplify = TRUE)
```

```{r}
x <- c("a", "a b", "a b c")
str_extract_all(x, "[a-z]", simplify = TRUE)
```

# Exercise
```{r}
# Matches "flickered" b/c it matches red
colour_match2 <- str_c("\\b(", str_c(colours, collapse = "|"), ")\\b")
colour_match2

more2 <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more2, colour_match2, match = TRUE)
```

# 14.4.3 Grouped matches
```{r}
noun <- "(a|the) ([^ ]+)"

has_noun <- sentences %>%
  str_subset(noun) %>%
  head(10)
has_noun %>% 
  str_extract(noun)
```

```{r}
has_noun %>% 
  str_match(noun)
# But smooth is not a noun? Doesnt really work
```

```{r}
tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)", 
    remove = FALSE
  )
```

# Exercise
```{r}
numword <- "\\b(one|two|three|four|five|six|seven|eight|nine|ten) +(\\w+)"
sentences[str_detect(sentences, numword)] %>%
  str_extract(numword)
```

# 14.4.4 Replacing matches
```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-") # Replaces only first instance of vowel
```

```{r}
str_replace_all(x, "[aeiou]", "-") # Replaces all
```

```{r}
x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```

```{r}
head(sentences)

sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)
```

# Exercise
```{r}
x <- "apples/bananas/pears"
str_replace_all(x, "/", "\\\\")
```

# 14.4.5 Splitting
```{r}
sentences %>%
  head(5) %>% 
  str_split(" ")
```

```{r}
"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]
```

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ", simplify = TRUE)
```

```{r}
fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(": ", n = 2, simplify = TRUE)
```

```{r}
x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))
```

```{r}
str_split(x, " ")[[1]]
```

```{r}
str_split(x, boundary("word"))[[1]]
```

# Exercise
```{r}
x <- c("apples, pears, and bananas")
str_split(x, ", +(and +)?")[[1]]
```

